### 2022-01-18 paper.Rmd
Update paper.Rmd

---
title: Investigating factors that may influence COVID-19 survival
author:
  - name: 36112985
    email: 36112985@lancaster.ac.uk
    affiliation: Lancaster Medical School
    corresponding: 36112985@lancaster.ac.uk
address:
  - code: Lancaster Medical School
    address: Lancaster Medical School, Faculty of Health & Medicine, University of Lancaster
abstract: |
  **OBJECTIVE** To investigate, amongst a set of probable prognostic factors, which factors may influence coronavirus 19 disease survival probabilities.  **DESIGN** Prospective cohort study; ongoing.  **SETTING** The International Severe Acute Respiratory and Emerging Infections Consortium (ISARIC) World Health Organization (WHO) Clinical Characterisation Protocol UK (CCP-UK) study, which enrols in-patients at at-least 260 Englsnd, Scotland, and Wales hospitals; the ISARIC Coronavirus Clinical Characterisation Consortium (ISARIC 4C) conducts the study. The investigation's patients are patients enrolled in a participating England hospital between 10 February 2020 and 5 July 2020.  PARTICIPANTS  Each study enrollee has a confirmed SARS-CoV-2 infection, or a high infection likelihood.  The investigation's patients are members of age groups *30 - 39*, *40 - 49*, and higher.  **OUTCOME** Patient survival.  **RESULTS** Inconclusive.  **CONCLUSIONS**  Insufficient spectrum of factors.

bibliography: paper/mybibfile.bib
output:
  rticles::plos_article:
    keep_tex: FALSE
# plos.csl can be download and used locally
csl: http://www.zotero.org/styles/plos
---

_Text based on plos sample manuscript, see <https://journals.plos.org/ploscompbiol/s/latex>_

# Introduction

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

A list

- Item 1
- Item 2

Here are two sample references: @Feynman1963118 [@Dirac1953888].

# Methods

## Exploratory Data Analysis

* Relationships
* Correlations

## Missing Data

Most of the data's variables have missing values.  Prior to deciding how to address missing values, it is important to understand the missing values patterns [@steyerberg2010]. Rubin [@rubin1976, @steyerberg2010, @little2019] outlines three fundamental missing data mechanisms

* missing completely at random (MCAR) / administrative errors, accidents
* missing at random (MAR) / missing data associated with known patient characteristics [independent variables], or the outcome
* missing not at random (MNAR) / missing data associated with missing values of the factor/predictor in question or with unobserved predictors

If the missing values of a data set in question are *missing completely at random* then complete case analysis will suffice because the complete case excerpt is akin to a random sample from a complete population.  If MCAR does not hold, e.g., data is *missing at random*, then the complete case excerpt is not representative of the underlying population, therefore population inference is not possible via complete case analysis. [@steyerberg2010]

Herein, missing data mechanisms analysis is via logistic regression.  Almost all the independent variables have missing values patterns that violate MCAR; addressed via multiple imputation in-line with the advice and protocols of [@sterne2009] & [@steyerberg2010].

# References {#references .unnumbered}